Xolotl: An Intuitive and Comprehensible Neuronal Simulator

### 2018-05-22 PaperLatex/frontiers.tex
Add new section to worked examples

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Hoyland, Gorur-Shandilya, \& Marder} %use et al only if is more than 1 author
\def\Authors{Alec Hoyland, Srinivas Gorur-Shandilya\,$^{*}$, and Eve Marder}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{Marder Lab, Brandeis University, Biology Department and Volen Center for Complex Systems, Waltham, MA, USA}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Srinivas Gorur-Shandilya
%
%Volen Center for Complex Systems
%
%Brandeis University
%
%415 South Street
%
%Waltham, MA 02454
%
%USA
}

\def\corrEmail{srinivasgs@brandeis.edu}

@@

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

@@

%%% Leave the Abstract empty if your article does not require one, please see the Summary Table for full details.
\texttt{xolotl} is an open-source neuronal simulator written in \texttt{C++} with \texttt{MATLAB} wrappers. Complex models and networks can be designed efficiently using an intuitive language tightly coupled to the object-based architecture of the underlying \texttt{C++} code. Models are specified by adding conductances to compartment objects. The structure is modular, serialized, and searchable, permitting high-level programmatic control over nearly all features of the models. \texttt{C++} templates are provided for developing new conductances, compartments, and integration schemata. \texttt{xolotl} readily implements parallel-processing on multicore processors and high-performance computing clusters. It also includes a customizable graphical user interface (`puppeteer`) for rapid prototyping and hand-tuning conductances in real-time. In addition, \texttt{xolotl} comes packaged with a powerful optimization toolbox (\texttt{procrustes}) for optimizing any model parameters accessible in the \texttt{xolotl} tree. The modular structure and accessibility to all parameters, variables, and dynamics of the model network in \texttt{MATLAB} facilitate rapid construction and assessment of model networks. \texttt{xolotl} is freely available at \url{https://github.com/sg-s/xolotl}. This tool provides straightforward implementation and fast simulation of neuronal models while permitting full control over every aspect of the network and integration.

@@

>>>>

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Hoyland, Gorur-Shandilya, \& Marder} %use et al only if is more than 1 author
\def\Authors{Alec Hoyland\,$^{\dagger}$, Srinivas Gorur-Shandilya\,$^{\dagger}$, and Eve Marder\,$^{*}$}

\def\Address{Marder Lab, Brandeis University, Biology Department and Volen Center for Complex Systems, Waltham, MA, USA}

\def\corrAuthor{Eve Marder

Volen Center for Complex Systems

Brandeis University

415 South Street

Waltham, MA 02454

USA
}

\def\corrEmail{marder@brandeis.edu}

@@

%\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
\extraAuth{$^{\dagger}$ Co-first authors}

@@

%%% Leave the Abstract empty if your article does not require one, please see the Summary Table for full details.
\texttt{xolotl} is an open-source neuronal simulator written in \texttt{C++} with \texttt{MATLAB} wrappers. Complex models and networks can be designed efficiently using an intuitive language tightly coupled to the object-based architecture of the underlying \texttt{C++} code. Models can be specified by adding conductances to compartment objects. The structure is modular, serialized, and searchable, permitting high-level programmatic control over nearly all features of the models. \texttt{C++} templates are provided for developing new conductances, compartments, and integration schemata. It also includes a customizable graphical user interface (`puppeteer`) for rapid prototyping and hand-tuning conductances in real-time. The modular structure and accessibility to all parameters, variables, and dynamics of the model network in \texttt{MATLAB} facilitate rapid construction and assessment of model networks. \texttt{xolotl} is freely available at \url{https://github.com/marderlab/xolotl}. This tool provides straightforward implementation and fast simulation of neuronal models while permitting full control over every aspect of the network and integration.

@@

\subsection{Simulating Integral Control}

\texttt{xolotl} can implement homeostatic tuning rules as integral control. Figure \ref{fig:figureintegralcontrol} depicts generation of a bursting neuron model from quiescent conditions. Calcium sensors supervene on maximal conductance density (\ref{fig:figureintegralcontrol}). In \texttt{xolotl}, integral controllers are properties of conductances (\ref{fig:figureintegralcontrol}B-C). They modify properties of the conductance in response to an error signal. In a demonstration adapted from O'Leary \textit{et al.} 2013, integral control changes maximal conductances to bring a neuron from quiescence into a bursting regime. Maximal conductances increase from random initial conditions to a set which elicits the desired network output by minimizing the error signal (\ref{fig:figureintegralcontrol}D-F).
